Program Mutation is Just Program Transformation: \\ A Universal Approach to Mutant Generation

### 2023-03-21 paper/main.tex
Alex first stab at intro

@@

\section{Experimental Evaluatoni}

>>>>

Mutation testing, though introduced in the late
1970s~\cite{demillo1978hints,mathur2012foundations,demillo1978hints},
has largely been a topic of academic interest rather than widespread
practice until quite recently.  However, with the adoption of mutation
testing by bellwether software industry companies including
Google~\cite{GoogleMut}, Meta~\cite{BellerFacebookMutation}, and
Amazon~\cite{AmazonMut}, interest in using mutation testing for
real-world projects has grown widely.

Interest in mutation testing has become widespread enough in the open
source community that a popular (more than 300 stars on GitHub)
repository, ``Awesome Mutation Testing,'' lists more than 40 tools,
covering almost twenty languages:
\url{https://github.com/theofidry/awesome-mutation-testing}.  Some of
these tools are the products of academic research, but many are
essentially the hobby projects of developers interested in enabling
mutation testing for their favorite language.

However, the same website demonstrates the existence of a sad history
of mutation testing tools: mutation tools have a tendency to be
developed for a while and then abandoned.   The page listing abandoned
mutation testing tools
(\url{https://github.com/theofidry/awesome-mutation-testing/blob/master/abandoned.md})
lists almost half as many tools as are under active maintenance, and
these tools cover approximately half as many languages.  Each of these
tools required considerable development time and effort, and was
popular enough at some point to be worth noting in the ``Awesome
Mutation Testing'' resource.  Many of the abandoned tools have GitHub
repositories (some archived) with more than 20 stars, and so may well
have had users who at some point faced the reality of no future bug
fixes or changes to address changes to the programming language being
mutated\footnote{While this is a minor concern for languages such as C
  that are relatively stable, it means that a mutation tool may fail
  to parse C++ 2020, and is particularly problematic for somewhat
  fast-moving languages such as Rust.}.  Of course, many open source
projects are abandoned, and academic software has a long history of
being available and supported only so long as a graduate student is
working on a thesis or a particular grant is being funded.  Still,
there appears to be a significant risk in relying on any mutation tools other
than the most widely used and well-supported, such as PIT.

...

In this paper, we argue that developing a large number of mutation
testing tools, each requiring its own development community, workflow
for mutation, and effort to keep up with language changes, is
unnecessary.  Mutation testing, and in particular, the core problem,
\emph{mutation generation} is simply a particular instance of the
problem of \emph{program transformations}.  Moreover,
mutation testing can be treated as an almost entirely \emph{syntactic}
transformation problem; in fact, in their classic introduction to
software testing, Ammann and Offutt refer to mutation testing as
``syntax-based testing'' ~\cite{ammann2016introduction}.  While, e.g., refactoring may need to
``understand'' a program to some extent, mutation can often operate
with essentially no context beyond a single line of code.

We further propose that in fact, for the most part, mutation
generation need not even ``know'' the syntax of a target language, in
terms of a complete grammar.  Instead, mutation testing can operate to
a large extent in a \emph{universal} manner, where programs are
transformed at the level of \emph{patterns of characters}.

The contributions of this paper are:

\begin{enumerate}
  \item The description of a novel approach to mutation generation
    based on generalized description of multi-lingual program transformations
    that allows a single tool to handle mutant generation for
    essentially all programming languages.  This approach also makes
    mutation testing easy to extend for new languages and even allows
    developers to write project-specific rules easily without
    having to modify a mutation tool's implementation.

          \item A proposal to use parser-parser combinators to improve the
      efficiency and expressive power of that approach, and an
      evaluation of the gains thus achieved.

      \item A comparison of an implementation of mutation generation
       based on this approach with existing tools for four important
       programming languages.  The single-language tools range from
       approximately 3,800 LOC to nearly 60,000 LOC, and each handle
       one programming language.  Our tool, which provides comparable
       core mutation testing functionality, supports over a dozen
       languages using only about 2200 LOC of Python and less than 500
       lines of rules defining mutation operators.

\end{enumerate}

@@

\section{Experimental Evaluation}

\begin{itemize}

  \item{RQ1:}  How does the use of parser-parser combinators modify 
    the efficiency (in terms of invalid mutants) and effectiveness (in
    terms of mutation score and equivalent mutants) of the universal
    approach to mutation generation?

\item{RQ2:}  How does the univeral approach, using regular expressions
  or parser-combinator combinator rules, compare to existing
  single-language mutation tools, in terms of number of generated
  mutants, mutation scores, equivalent mutants, and other evaluation measures used in the literature?

  \end{itemize}
